---
title: "mean-centering"
author: "rccline"
date: "6/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Mean Centering by Shawn Janzen
* 13.10 Multiple Linear Regression: Mean-Center & Standardization 
* [Shawn Janzen's YouTube channel](https://www.youtube.com/watch?v=FYjuu9_5oMQ) 

   


1. Take all of OBS and subract the mean -> And the new average is zero.
2. Standardization does the same but obs/sd
    will transform the mean centered scores into z scored.

3.  instead of being spread in terms of tne measured variable,
     it will be spread in terms of the standard deviation,
     i.e. "normalized distances from the mean"

Why do this?
4. Both have a mean of zero
5. The mean-centered can interpret a y-intercept 
    (standardized, not so much)
6. mean centered, retains the unit of measure
7. mean centered Scales stays the same = same slope coefficients
8. Standardized, Rescales the variable to SD
9. Standardized Can help with multicollinearity
10. Standadrized can help When you have interactions, or polynomial terms
. 

mtcars$hp vector
```{r}
mtcars$hp
```


## Mean-Center----
```{r}
mtcars$hp.mc <- mtcars$hp - mean(mtcars$hp)
```



## Standardize to z-scores----
*scale is a base R function
```{r}
mtcars$hp.st <- scale(mtcars$hp)
```



## Look at Regression coefficients----
```{r}
summary(lm(qsec ~ hp, data=mtcars))

summary(lm(qsec ~ hp.mc, data=mtcars))

summary(lm(qsec ~ hp.st, data=mtcars)) 
``` 


### Custom function:  *ggplotRegression*
* [Code from Susan Johston, https://sejohnston.com/ Royal Society University Research Fellow, - Johnson Lab](https://sejohnston.com/2012/08/09/a-quick-and-easy-function-to-plot-lm-results-in-r/)
```{r}
library(tidyverse)
ggplotRegression <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     "Intercept =",signif(fit$coef[[1]],5 ),
                     " Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5)))
} 
```



```{r eval=FALSE}
library(tidyverse)


ggplotRegression <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     "Intercept =",signif(fit$coef[[1]],5 ),
                     " Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5)))
} 
```


#### lm() plot qsec ~ hp
```{r fig.cap='Fig.1 qsec ~ hp, Original data, not centered.', fig.topcaption=TRUE}
ggplotRegression(lm(qsec ~ hp, data=mtcars))

``` 

#### lm() plot qsec ~ hp.mc
```{r  fig.cap='Fig.2 qsec~hp.mc Mean-Centered.', fig.topcaption=TRUE}
ggplotRegression(lm(qsec ~ hp.mc, data=mtcars))

```


#### lm() plot qsec ~ hp.st
```{r fig.cap='Fig.3 qsec~hp.st SD-Centered.', fig.topcaption=TRUE}
ggplotRegression(lm(qsec ~ hp.st, data=mtcars))
```
 
 
#### Summary Tables (c.f. Lüdecke, 2021) 
```{r} 

# https://cran.r-project.org/web/packages/sjPlot/vignettes/tab_model_estimates.html 
# load package
library(sjPlot)
library(sjmisc)
library(sjlabelled)  


```

```{r}
hp1 <- lm(qsec ~ hp, data=mtcars)
hp2 <- lm(qsec ~ hp.mc, data=mtcars)
hp3 <- lm(qsec ~ hp.st, data=mtcars)
```

```{r}
tab_model(hp1, hp2, hp3)
#tab_model(hp2)
# tab_model(hp3)
```

Lüdecke, D. (2021, May 25). Data Visualization for Statistics in Social Science [R package sjPlot version 2.8.8]. Comprehensive R Archive Network (CRAN). https://CRAN.R-project.org/package=sjPlot

